
\subsection{Objectives}
The following objectives should be solved in my final year project:
\begin{enumerate}
\item Constructing an experimental environment for users of Genetic Algorithms.\\
Learning the use of Genetic Algorithms is very complicated for newcomers in this
area. Aim in this project is the construction of a number of libraries which
will help a potential user to construct genetic systems.
\item The environment should be programmed in Ada.\\
Basic idea is that the programming should be in an object orientated language which could be used
by a greater number of users. The computer language, which is best known by the
students and also most portable between different system is Ada. An alternative
would be the language C++, but there are not enough platforms availiable.
\item The packages should represent the different levels of Genetic Algorithms.\\
A first package cover simple Genetic Algorithms as described by Goldberg\cite{Gol89}.
econd package is an extension of the first also covering the advanced mechanims.
A third package is the implementation for two dimensional individuals.
\item The report should help a newcomer to learn the use of Genetic Algorithms.\\
The report has two parts. The first part is a short introduction in the theory
of Genetic Algorithms, the second part is a case study on an existing problem.
\end{enumerate}
The packages will be available on three platforms :
\begin{itemize}
\item on the Apollo Workstations
\item on MS-Dos Computers with an Ada compiler
\item on Unix System V/3.2 with Meridian Ada compiler
\end{itemize}

The report is written with the \TeX \  wordprocessor, using the \LaTeX \  macro package.
