\documentstyle[titlepage]{report}
%macropackage=lplaing

\author{{\sc Andreas Grimm}\\
{\small Dipl.-Inform., B.Sc.(Hons)}}

\def\bs{\char'134 } % backslash in \tt font.
\title{\sc Development of an Experimental
           System for the use of\\
           Genetic Algorithms}
\frenchspacing
\sloppy
\date{\today}
\pagestyle{headings}
\parindent0em \parskip1.5ex plus0.5ex minus 0.5ex
\renewcommand{\baselinestretch}{1.5}
\begin{document}
\newtheorem{definition}{Definition}[chapter]
\newtheorem{satz}{Satz}[chapter]
\newtheorem{hypothese}{Hypothese}[chapter]
\maketitle
\begin{abstract}
The following report is a reviewed verison of a report which I wrote in 1992 as
a Project Report for the University of Sunderland. 
It consists of five parts. The first part is a short introduction
into the area of genetic algorithms. It shows the fundamentals of the theory
of genetic algorithms and explains most of the terminology. Also a short example
is given, so that the elementary work of the genetic algorithms is explained.
The second chapter shows how the problem of the first chapter can be programmed.
The third chapter expands the theory and try to explain certain mechanisms which
are developed in the last years. An example of these advanced mechanisms is shown
in the fourth chapter. The last part is an introduction into the area of two
dimensional genetic algorithms and a short discussion of the possibilities for
the use of these.

The appendix gives a documentation for the used packages and introduces an
optional random number generator, which could be used independently from this
work.
\end{abstract}
\pagenumbering{Roman}
\tableofcontents
\newpage
\pagenumbering{arabic}
\section{Introduction}
\input introduc
\section{Acknowledgements}
At this part I want to thank my superviser, Dr. Peter Dunne for his support,
especially for his help by writting this report.
\chapter{The Basics of Genetic Algorithms}
\input chapter1
\chapter{Programming a SGA}
\input chapter2
\chapter{Advanced Genetic Algorithms}
\input chapter3
\chapter{Programming with AGA's}
\input chapter4
\chapter{Two Dimensional GAs}
\input chapter5
\begin{appendix}
\chapter{The Package for the SGA}
\input anhang_a
\chapter{The Package for the AGA}
\input anhang_b
\chapter{The Random Number Generator}
\input anhang_d
\input bibgraph
\end{appendix}
\end {document}
